\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-04-08 honors_thesis/journal_paper/paper.tex
rewrote my least-favorite paragraph, but is it now better or worse...

In this form the important point to notice is that there are only 7 values (for
this Ring topology) in the set formed by $\specpos_{t+1}^{c} \ c \in\ \caseset$
and that none of them depend upon $f(\ppos_t)$ or any other objective function
evaluation at iteration $t$ (note also that while there are random numbers in
the equation, they are assumed fixed once drawn for any particular particle at
a specific iteration).  Thus PSO has been refactored such that the algorithm
can begin computing all 7 of the objective function evaluations potentially
needed in iteration $t+1$ \emph{before} $f(\ppos_t)$ is computed.  Once the
evaluation of $f(\ppos_{t})$ is completed for all particles only one of the
indicator functions $\indic_{t+1}^{c}$ will be set to 1.  We will refer to
the indicator function which is set to 1 as the ``behavior preserving
indicator'' or $bp$, since it indicates which of the computations
$\ofunc(\specpos_{t+1}^{c})$ preserves the behavior of PSO.  The rest of the
computations $\ofunc(\specpos_{t+1}^{c}) \ c \in \caseset-\{bp\}$ will be
ignored, and might just as well never have been computed.  We call these
evaluations of $\ofunc(\specpos_{t+1}^{c})$ ``speculative children'' because
they may or may not be needed.

@@

The value of the refactoring can best be seen in a small example.  Suppose that
unlimited processors were available and that the evaluation of the objective
function took one hour. Suppose also that for each particle the computations
$\ofunc(\ppos_{t})$ and all of the $\ofunc(\specpos_{t+1}^{c})$ were sent to
separate processors and completed in parallel.  Once the values of
$\ofunc(\ppos_{t})$ have been computed for all particles, the computation
needed to determine the values of the indicator functions $\indic_{t+1}^{c}$
and to complete iteration $t+1$ is inconsequential.  In this environment all of
the function evaluations needed for two iterations of PSO could be completed in
just one hour where as the conventional implementation would require two hours
for those same two iterations.

>>>>

In this form the important point to notice is that there are only 7 values
(for this Ring topology) in the set $\{\specpos_{t+1}^{\casegen}: \casegen \in
\caseset\}$ and that none of them depend upon $f(\ppos_t)$ or any other
objective function evaluation at iteration $t$ (note also that while there are
random numbers in the equation, they are assumed fixed once drawn for any
particular particle at a specific iteration).  Thus PSO has been refactored
such that the algorithm can begin computing all 7 of the objective function
evaluations potentially needed in iteration $t+1$ \emph{before} $f(\ppos_t)$
is computed.  Once the evaluation of $f(\ppos_{t})$ is completed for all
particles only one of the indicator functions $\indic_{t+1}^{\casegen}$ will
be set to 1.  Although this speculative algorithm computes
$\ofunc(\specpos_{t+1}^{\casegen})$ for all $\casegen \in \caseset$, even
those for which $\indic_{t+1}^{\casegen} = 0$, these extra computations will
be ignored, and might just as well never have been computed.  We call the set
of computations $\{\ofunc(\specpos_{t+1}^{c}) : \casegen \in \caseset\}$
``speculative children'' because only one of them will be needed.

@@

Suppose that unlimited processors were available and that the evaluation of
the objective function took one hour. Suppose also that for each particle the
computations $\ofunc(\ppos_{t})$ and all of the $\ofunc(\specpos_{t+1}^{c})$
were sent to separate processors and completed in parallel.  Once the values
of $\ofunc(\ppos_{t})$ have been computed for all particles, the computation
needed to determine the values of the indicator functions $\indic_{t+1}^{c}$
and to complete iteration $t+1$ is inconsequential.  In this environment all
of the function evaluations needed for two iterations of PSO could be
completed in just one hour where as the conventional implementation would
require two hours for those same two iterations.
